---
title: "Project 3 - Fitness Tracking"
output: html_document
---

```{r setup, include=FALSE}

#Basic set-up to load libraries and project data. 
library(dplyr)
library(stringr)
library(ggplot2)
library(tidylog)
library(tidyr)
library(janitor)
library(digest)
library(purrr)
library(lubridate)
library(styler)

library(readr)
fitness_members <- read_csv("~/Desktop/EPFL Data Science Course/fitness_members.csv")
fitness_tracking <- read_csv("~/Desktop/EPFL Data Science Course/fitness_tracking.csv")

#Column names are already formatted okay so no clean up is needed. 

```

#### Part 1. 
**Basic breakdown of gym members.**
```{r basic_view, echo=FALSE, message=FALSE}
fitness_members %>% 
  tabyl(gender) %>% 
  adorn_pct_formatting(digits = 1)
```

#### Part 2: 
**How many successful recommendations has member 000115 made?**
```{r recommendations_part2, echo=FALSE, message=FALSE}
number_recommendations_000115 <- 
  fitness_members %>% 
  filter(recommendation_from =="000115") %>% 
  print()
```

#### Part 3: 
**What is the maximum number of successful recommendations made by a member of the Fitness Club?**
#### Part 4: 
**What is the id code of the member who made the maximum number of successful recommendations?**
```{r recommendations_parts34, echo=FALSE, message=FALSE}
number_recommendations <- 
  fitness_members %>% 
  filter(recommendation_from !="NA") %>% 
  arrange(recommendation_from) %>% 
  tabyl(recommendation_from) %>% 
  arrange(desc(n)) %>% 
  head(5)

number_recommendations
```

#### Part 5. 
**How many female members in the Premium membership category made one or more successful recommendations?**
```{r female_premiums, echo=FALSE, message=FALSE}
#Option no1 to achieve this result (a slightly convoluted way)
ids_of_female_premiums <- 
 fitness_members %>% 
   filter(gender == "F", m_category == "Premium") %>% #Filter our Female Premium members
   select(recommendation_from = id)  #Rename id column to recommendation_from to compare the lists with the inner join.

inner_join(ids_of_female_premiums, fitness_members, by = c("recommendation_from"))

#Option no2 (and a simpler way) to achieve the same result.
ids_of_female_premiums <- 
 fitness_members %>% 
  filter(gender == "F", m_category == "Premium") %>% #Filter our Female Premium members
  pull(id)

fitness_members %>% 
 filter(recommendation_from %in% ids_of_female_premiums)
```

#### Part 6: What is the BMI value of member 000042 at registration date?  
(Note:BMI is calculated weight/height^2)
```{r BMI, echo=FALSE, message=FALSE}
#Convert height values from cm into m
BMI_000042 <-
fitness_members %>% 
  mutate(height = height/100, BMI = weight/(height^2)) %>% #Convert cm to m for height and add a column for BMI.
  filter(id == "000042") %>% #Display result only for member 000042
  pull(BMI)


```
The BMI of member 000042 at registration was `r round(BMI_000042, digits=2)`

#### Part 7. 
As mentioned in the introduction, the Fitness Club has 420 members. Some of them have not yet completed the first week. Those members are not included in the fitness_tracking.csv data file.

Which fitness club members have not yet reached week 1?
```{r week_1, echo=FALSE, message=FALSE}
fitness_members %>% 
  full_join(fitness_tracking, by = c("id")) %>%  #Join the two tibbles together, retaining all rows.
  filter(is.na(wk_001)) %>% #Show only rows where there is no value yet for wk_001, i.e. members that have not yet completed week 1.
  select(id) 

```

#### Part 8. 
Convert member weights into a long table (in preparation for Project 4).  
Note: Discard from fitness_tracking_long all the rows where no weight value is available. And make sure that wk_000 values (i.e. values at registration date) are also included.
```{r tracking_long, message=FALSE, warning=FALSE, include=FALSE}
fitness_tracking_long <-
fitness_members %>% 
  full_join(fitness_tracking, by = c("id")) %>% 
  mutate(wk_000 = weight) %>%                   #Include also joining weight as wk_000
  select(1, wk_000, starts_with("wk")) %>%      #Exclude all columns apart from id and weight related colums
  pivot_longer(
    cols= c(starts_with("wk")),
    names_to = "week", 
    values_to = "weight") %>% 
  filter(weight != "")   #Exclude rows where no weight value is available.

```

#### Part 9. Use the data set fitness_tracking_long to compute weekly BMI values for each member. 

**What is the BMI value of member 000024 at week 11?**
```{r weekly_BMI, echo=FALSE, message=FALSE, warning=FALSE}
fitness_tracking_long_bmi <-
fitness_members %>% 
  full_join(fitness_tracking, by = c("id")) %>% 
  mutate(wk_000 = weight) %>%                   #Include also joining weight as wk_000
  select(id, height, wk_000, starts_with("wk")) %>%      #Exclude all columns apart from id and weight related colums
  group_by(id) %>% 
  pivot_longer(
    cols= c(starts_with("wk")),
    names_to = "week", 
    values_to = "weight") %>% 
  filter(weight != "") 

BMI_000024_11 <-
fitness_tracking_long_bmi %>% 
  mutate(height = height/100, BMI = weight/(height^2)) %>% 
  filter(id == "000024", week == "wk_011") %>% 
  pull(BMI)
```
The BMI of member 000024 is `r round(BMI_000024_11, digits=2)`

#### Part 10.  
**What is the BMI percentage change for member 000015 from week 6 to week 7?**
```{r BMI_change, include=FALSE}
perc_chage_000015 <-
fitness_tracking_long_bmi %>% 
  mutate(height = height/100, BMI = weight/(height^2), BMI_change = (BMI/lag(BMI, n=1)-1)) %>% 
  filter(id=="000015", week == "wk_007") %>% 
  pull(BMI_change)
```
The user experimented a drop in BMI of  `r round(perc_chage_000015*100, digits=2)` %
